---
title: "What to forecast"
editor: 
  markdown: 
    wrap: sentence
---

## Targets file

Daily time-step variables measured within the reservoirs:

```{r}
url <- "https://renc.osn.xsede.org/bio230121-bucket01/vera4cast/targets/duration=P1D/daily-insitu-targets.csv.gz"
```

```{r message=FALSE}
library(tidyverse)
targets <- read_csv(url, show_col_types = FALSE)
```

```{r}
glimpse(targets)
```

### Variable list 

```{r message=FALSE}
googlesheets4::gs4_deauth()
target_metadata <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1fOWo6zlcWA8F6PmRS9AD6n1pf-dTWSsmGKNpaX3yHNE/edit?usp=sharing")
```

```{r echo = FALSE}
  target_metadata <- target_metadata |> 
  rename(variable = `"official" targets name`,
         priority = `priority target`) |> 
  select(variable, duration, class, Description, priority)
```

#### Physical variables

```{r echo = FALSE}
targets |> 
  distinct(variable, duration) |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  filter(class == "physical" & priority) |> 
  select(-class,-priority) |> 
  knitr::kable()
```
```{r echo = FALSE, warning = FALSE}
targets |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  filter(class == "physical" & priority) |> 
  select(-class,-priority) |> 
  filter(depth_m %in% c(1.5, 1.6) | is.na(depth_m)) |> 
  ggplot(aes(x = datetime, y = observation)) +
  geom_point() +
  facet_grid(variable~site_id, scales = "free_y") +
  theme_bw()
```
#### Chemical variables

```{r echo = FALSE}
targets |> 
  distinct(variable, duration) |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  filter(class == "chemical" & priority ) |> 
  select(-class,-priority) |> 
  knitr::kable()
```
```{r echo = FALSE, warning = FALSE}
targets |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  filter(class == "chemical" & priority) |> 
  select(-class,-priority) |> 
  filter(depth_m %in% c(1.5, 1.6) | is.na(depth_m)) |> 
  ggplot(aes(x = datetime, y = observation)) +
  geom_point() +
  facet_grid(variable~site_id, scales = "free_y") +
  theme_bw()
```
#### Biological variables

```{r echo = FALSE}
targets |> 
  distinct(variable, duration) |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  filter(class == "biological"& priority) |> 
  select(-class,-priority) |> 
  knitr::kable()
```
```{r echo = FALSE, warning = FALSE}
targets |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  filter(class == "biological" & priority) |> 
  select(-class,-priority) |> 
  filter(depth_m %in% c(1.5, 1.6) | is.na(depth_m)) |> 
  ggplot(aes(x = datetime, y = observation)) +
  geom_point() +
  facet_grid(variable~site_id, scales = "free_y") +
  theme_bw()

```

## Site

```{r}
site_list <- read_csv("https://raw.githubusercontent.com/LTREB-reservoirs/vera4cast/main/vera4cast_field_site_metadata.csv", show_col_types = FALSE)
```

```{r echo = FALSE}
site_list |> knitr::kable()
```

## Additional target variables

### Stream inflow to Falling Creek Reservoir

Daily time-step variables measured in the monitored stream (Tunnel Branch; site_id = tubr)

```{r}
url2 <- "https://renc.osn.xsede.org/bio230121-bucket01/vera4cast/targets/duration=P1D/daily-inflow-targets.csv.gz"
inflow_targets <- read_csv(url2, show_col_types = FALSE)
glimpse(inflow_targets)
```
```{r echo = FALSE}
inflow_targets |> 
  distinct(variable, duration) |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  select(-class) |> 
  knitr::kable()
```

### Falling Creek Reservoir meteorology

Daily time-step variables measured at the meteorology station (located on the dam at Falling Creek Reservoir)

```{r}
url3 <- "https://renc.osn.xsede.org/bio230121-bucket01/vera4cast/targets/duration=P1D/daily-met-targets.csv.gz"
met_targets <- read_csv(url3, show_col_types = FALSE)
glimpse(met_targets)
```

```{r echo = FALSE}
met_targets |> 
  distinct(variable, duration) |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  select(-class) |> 
  knitr::kable()
```

Hourly time-step variables measured at the meteorology station (located on the dam at Falling Creek Reservoir)

```{r}
url3 <- "https://renc.osn.xsede.org/bio230121-bucket01/vera4cast/targets/duration=PT1H/hourly-met-targets.csv.gz"
```

### Other variables measured in the reservoirs

```{r echo = FALSE}
targets |> 
  distinct(variable, duration) |> 
  left_join(target_metadata, by = c("variable","duration")) |> 
  filter(!priority) |> 
  select(-class,-priority) |> 
  knitr::kable()
```

